---
title: "My first markdown file"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(ggplot2)
```


## Reading the Data

The fav_restaurants dataset involves a csv file containing names of the students, along with their favorite restaurants, the location of said restaurants, and the average prices for each meal. The data set also contains the last time visited in the form "YYYY-MM".

The below code loads the csv file:
```{r }
fav_restaurants <- read_csv("C:\\Users\\ug5fw\\Desktop\\fav_restaurants.csv")
```


## Plotting the Data

This plot represents the a bar graph of all submissions based on type of cuisine. The x axis is the type of cuisine while the y axis is the amount of each type.
```{r}
ggplot(fav_restaurants, aes(cuisine)) + geom_bar()
```




This plot represents a point graph off all submissions based on the cost per person based off of the cuisine type.
```{r}
ggplot(fav_restaurants, aes(cuisine, cost_per_person )) + geom_point()
```




This point plot represents all submissions based on the name of the restaurant with yearly visits.
```{r}
ggplot(fav_restaurants, aes(name, yearly_visits)) + geom_point()
```