---
title: "Lab 18: Inference for the Mean"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Teacher Salary

We will start today by looking at a small dataset containing teacher
salaries from 2009-2010 for 71 randomly choosen teachers employed by the
St. Louis Public School in in Michigan.

```{r}
teachers <- read_csv("https://statsmaths.github.io/stat_data/teachers_pay.csv")
```

The available variables are

- base: Base annual salary, in dollars
- degree: Highest educational degree attained: BA (Bachelor's) or MA (Master's)
- years: Number of years employeed

Using the mean function, what is the average base pay of all teachers
in the dataset? 56937.61 dollars

```{r}
base_pay_mean <- mean(teachers$base)
base_pay_mean
```

Fit a model for the mean of the base pay variable using lm_basic. Save
the model as an object called "model":

```{r}
model <- lm_basic(base ~ 1, data = teachers)
model
```

Using a call to `reg_table`, find the mean implied by the model:

```{r}
reg_table(model)
```

Does the mean agree with your answer to question 2? 

**Answer**: Yes. The differences between the two are negligible.

Add a 95% confidence interval to the regression table.

```{r}
reg_table(model, level = 0.95)
```

What is the range of mean salaries implied by the confidence interval?

**Answer**: Anywhere from 54785(2.5%) to 59091(97.5%) dollars

Draw a histogram of the base salary values for the entire dataset.

```{r}
ggplot(teachers, aes(base)) +
  geom_histogram()
```

Do most of the salary values fall within the range given in question 5? Why
or why not?

**Answer**: Yes because the approximate total count for the prices in the range from ~54000 to ~59000 is larger than the ones before and after this range.

Use the filter command to construct a new dataset called `masters` consisting
of just those teachers with a masters degree.

```{r}
masters <- filter(teachers, degree == 'MA')
masters
```

Compute a 95% confidence interval for the mean pay of teachers with a master's
degree. Does this range intersect the one you had in question 5? Yes

```{r}
reg_table(lm_basic(base ~ 1, data = masters), level = 0.95)
```

## Murder Data

Now load the following dataset containing all murders that have occurred in
London from 1 January 2006 to 7 September 7 2011.

```{r}
london <- read_csv("https://statsmaths.github.io/stat_data/london_murders.csv")
```

The available variables are:

- age: age of the victim in years
- year: year of the murder
- borough: the London borough in which the murder took place

Find an 80% confidence interval for the average age of the victim of a
murder in London.

```{r}
reg_table(lm_basic(age ~ 1, data = london), level = 0.8)
```

Make sure you actually extract the answer here:

**Answer**: Average age -> 33.86 years

Describe in words how the proceeding confidence interval should be
interpreted.

**Answer**: Best guess is 33.86 years, but it is worth mentioning that murders in the range from 33.07 to 34.65 should be acceptable.
