---
title: "Lab 22: Sixteen Personality Factors II"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## 16PF Data, Again

Once again we will load the survey response data from the Sixteen Personality
Factor Questionnaire in order to practice our skills at statistical inference:

```{r}
pf <- read_csv("https://statsmaths.github.io/stat_data/cattell_16pf.csv")
```

The dataset use the following fields:
- age: respondent's age in years
- gender: respondent's self-selected gender
- country: two letter IATA code for the respondent's IP
- elapsed: time taken to complete quiz in seconds
- warmth: personality score from 1-20
- reasoning: personality score from 1-20
- emotional_stability: personality score from 1-20
- dominance: personality score from 1-20
- liveliness: personality score from 1-20
- rule_consciousness: personality score from 1-20
- social_boldness: personality score from 1-20
- sensitivity: personality score from 1-20
- vigilance: personality score from 1-20
- abstractedness: personality score from 1-20
- privateness: personality score from 1-20
- apprehension: personality score from 1-20
- openness_to_change: personality score from 1-20
- self_reliance: personality score from 1-20
- perfectionism: personality score from 1-20
- tension: personality score from 1-20
- baseline: average score across all 16 personality traits

Use the `percentiles` function to detect the 3th and 97th percentiles for
the amount of elapsed time taken to complete the quiz.

```{r}
percentiles(pf$elapsed)
```

**Answer**: 3% = 369s. 97% = 4107.46s

We want to construct a new variable called speed that tells us whether
someone was a fast test taker (less than the 3rd percentile), a slow test
taker (greater than the 97th percentile), or a normal test taker. To do
so, I'll use some code that we have not seen in this form before.
If the 3rd percentile was 100s and the 97th was 1000s, it would look like
this:

```{r}
pf$speed <- "normal"
pf$speed[pf$elapsed < 369.00] <- "fast"
pf$speed[pf$elapsed > 4107.46] <- "slow"
```

Modify the code above use the cutoffs you found in the previous question (and
make sure you run it).

Fit a linear model predicting the prefectionism variable as a function
of the speed variable.

```{r}
model <- lm_basic(perfectionism ~ 1 + speed, data = pf)
reg_table(model, level = 0.95)
```


How does the perfectionism score differ between the three groups? Are both
slopes significant?

**Answer**: Both slopes are statistically significant, and there is not much of a difference between the normal and slow slopes.

In the last question it should have appeared that the perfectionism
score for the slow and normal groups were very similar. In the regression
as given there is no way to test whether these two groups have a
statistically significant difference because the slopes only relate to the
baseline level (here, "fast").

It is possible to change the baseline level of a categorical variable in
R. To do so, use the function `fct_relevel`; its second argument gives the
desired baseline level:

```{r}
pf$speed_normal <- fct_relevel(pf$speed, "normal")
model <- lm_basic(perfectionism ~ 1 + speed_normal, data = pf)
reg_table(model, level = 0.95)
```

Using this table, is there a statistically significant difference between
normal and slow users in their perfectionism score?

**Answer**: the slow slope is not statistically significant since it passes through zero. However, there is a statistically significant difference between fast and slow since the difference between the two slopes is very large.

Based on the result in the previous describe a
plausible reason and a some people may finish quickly and a plausible reason
some may take a long time. Hint: Converting the 97th percentile to hours
may give you some ideas for the last part.

**Answer**: Some people may already know themselves really well, in which case answering the questions might be easier. Otherwise, some people may just choose random answers and see what trait they get.

## Selected trait, again

Take your personality trait from last class and select the one trait
that most closely contrasts with your trait and one that most closely
matches your trait. Fit a linear regression model that predicts your trait
as a function of both of these as well as the baseline variable.

```{r}
model <- lm_basic(self_reliance ~ 1 + social_boldness + emotional_stability, data = pf)
reg_table(model, level = 0.95)
```

Interpret the significance and signs of the slopes in the previous model
for the two traits that you selected. Do their signs match what you would
have expected?

**Answer**: Both social boldness and emotional stability are statistically significant. The signs do match what I expected since people who are more emotionally stable could tend to prefer to have time with themselves since they know themselves the best. For social boldness I would be positive since people that are more self reliant could prefer to spend most of their time in solitude.

Take the previous model and add the variables `speed`, `gender`, and
`country`, the latter lumped into 5 categories.

```{r}
grouped_countries <- fct_lump(pf$country, n=5)
model <- lm_basic(self_reliance ~ 1 + social_boldness + emotional_stability + speed + gender + grouped_countries, data = pf)
reg_table(model, level = 0.95)
```

Does this change the slopes for your two traits much?

**Answer**: No

Summarize the previous model (specifically the relationship between the
three traits) in words.

**Answer**: People who showcase higher levels of self-reliance tend to be better emotionally stable and more socially able. There isn't much correlation in terms of test taking speeds for people who are more self-reliant. Males tend to be less self-reliant as well. For country regions, most countries except in canada tend to show less self-reliance.

