---
title: "Lab 24: Chicago Crime Data I"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Chicago Crime Data

Today we are going to look at a fairly largerdataset. Each row
of the data refers to a single reported crime in the City of Chicago:

```{r}
crimes <- read_csv("https://statsmaths.github.io/stat_data/chi_crimes_2016.csv")
```

The available variable are:

- `area_number`: the community area code of the crime; a number from 1-77
- `arrest_flag`: whether the crime resulted in an arrest; 0 is false and 1 is true                
- `domestic_flag`: whether the crime is classified as a domestic offense; 0 is false and 1 is true                    
- `night_flag`: did the crime occur at night (9pm - 3am); 0 is false and 1 is true               
- `burglary`: was the crime classified as a burglary? 0 is false and 1 is true
- `theft`: was the crime classified as a theft? 0 is false and 1 is true
- `battery`: was the crime classified as a battery? 0 is false and 1 is true
- `damage`: was the crime classified as a damage? 0 is false and 1 is true
- `assault`: was the crime classified as an assault? 0 is false and 1 is true
- `deception`: was the crime classified as criminal deception? 0 is false and 1 is true
- `robbery`: was the crime classified as a robbery? 0 is false and 1 is true
- `narcotics`: was the crime classified as a narcotics violation? 0 is false and 1 is true

We also have metadata about each community area within Chicago as well.
We will see how to use these shortly.

```{r}
ca <- read_csv("https://statsmaths.github.io/stat_data/chicago_meta.csv")
```

- `area_number`: the community area code; a number from 1 to 77
- `area_name`: popular name of the community area
- `median_age`: the median age of all residents in the community area
- `num_households`: total number of households
- `family_households`: percentage of households classified as a `family'
                     (domestic partners, married couples, and one or more
                     parents with children)
- `family_w_kids`: percentage of households with children under the age of 18                     
- `owner_ratio`: ratio of households that own or mortgage their primary residence                    
- `mean_travel_time`: average commute time
- `percent_walk`: percentage of commuters who walk to work (0-100)
- `median_income`: median household income
- `perc_20_units`: percentage of residential buildings with 20 or more units


It is difficult to do much of anything directly with the raw data. We
need to utilize the group_summarize function to get somewhere interesting.
Before doing that on the whole dataset, let's make sure that we understand
exactly what is going on by using the mean() and sum() functions directly.
Take the mean of arrest_flag for the whole dataset:

```{r}
mean(crimes$arrest_flag)
```

Describe what this means in words.

**Answer**: Around 17.2% of crimes resulted in an arrest.

Now, take the mean of the theft variable over the entire dataset:

```{r}
mean(crimes$theft)
```

Describe what this means in words.

**Answer**: Around 27.5% of all crimes committed were thefts.

Take the sum of the theft variable:

```{r}
sum(crimes$theft)
```

Describe what this means in words.

**Answer**: Out of 141064 total crimes committed, 38816 were thefts.

Take the dataset `ca` and calculate the sum of the variable `num_households`:

```{r}
sum(ca$num_households)
```

Divide the sum of the theft variable by the sum of the number of
households variable and multiply by 1000.

```{r}
(sum(crimes$theft) / sum(ca$num_households)) * 1000
```

Describe what this means in words.

**Answer**: 36.5% of Chicago households were affected by theft crimes.

Use the filter function to construct a dataset `temp` consisting only of
those rows in `crimes` that come from area_number 23. This is the area
named Humboldt Park.

```{r}
temp <- filter(crimes, area_number == "23")
```

Take the mean of the variable arrests on the data `temp`.

```{r}
mean(temp$arrest_flag)
```

Is this smaller, larger, or about the same as the mean of the arrest flag
over the entire dataset? Can we safely compare these measurements? If so
describe the relationship in words.

**Answer**: Larger than the mean of the arrest flag over the entire dataset. Most crimes that result in arrest could be concentrated around area number 23.

Take the mean of the variable `theft` on the data `temp`.

```{r}
mean(temp$theft)
```

Is this smaller, larger, or about the same as the mean of the theft flag
over the entire dataset? Can we safely compare these measurements? If so
describe the relationship in words.

**Answer**: Smaller than the mean of the theft flag over the entire dataset. Less thefts tend to happen around area number 23 compared to the rest of the city. 

Manually look up the number of households in area 23, Humboldt Park, by
looking at the dataset `ca` in the data viewer. Take the sum of the number
of thefts in temp, divide this by the number of households in Humboldt Park,
and multiply by 1000:

```{r}
(sum(temp$theft) / 17830) * 1000
```

Is this smaller, larger, or about the same as the same measurement over the
over the entire dataset? Can we safely compare these measurements? If so
describe the relationship in words.

**Answer**: It is larger then the same measurement over the entire dataset. Humboldt Park experiences more thefts when compared to the rest of the city.

You have the tools to get the number of households in Humboldt Park
entirely using R code. Specifically, filter `ca` to only include area 23
and then pull off `num_households` with the `$` operator. Try this here:

```{r}
new_temp <- filter(ca, area_number == "23")
humboldt <- new_temp$num_households
humboldt
```

Now that we have an idea of what it means to take means and sums over
this dataset, use `group_summarize` to summarize `crimes` at the community
area level. Save the result as `crimes_ca`:

```{r}
crimes_ca <- group_summarize(crimes, area_number)
```

We want to combine the datasets `crimes_ca` and `ca`. To do this we use
a new function `left_join`, as follows:

```{r}
crimes_ca <- left_join(crimes_ca, ca)
```

What variable is R using to match these datasets up?

**Answer**: area_number

By default, R will use any commonly named variables to match the datasets
up. If we need it, I will show you how to modify this behavior later.

Construct a variable `theft_rate` equal to the number of thefts in each
community area, divided by the number of households and multiplied by 1000.

```{r}
crimes_ca$theft_rate <- (crimes_ca$theft_sum / crimes_ca$num_households) * 1000
```

Draw a scatter plot with median income on the x-axis and theft rate on
the y-axis.

```{r}
ggplot(crimes_ca, aes(median_income, theft_rate)) + 
  geom_point()
```

Describe the general pattern and any outliers.

**Answer**: As median income increases for a neighborhood, the theft rate decreases. There is one outlier, an area where theft rate reaches well over 300l.

Modify the previous plot so that the points are labeled with the area
name:

```{r}
ggplot(crimes_ca, aes(median_income, theft_rate)) + 
  geom_point() +
  geom_text(aes(label = area_name))
```

What is the name of the outlier?

**Answer**: Loop

(Critical Thinking) Look up the neighborhood that is the outlier on
Wikipedia. Give an explanation for why this neighborhood has a high rate
of theft per household. Hint: There are two distinct reasons.

**Answer**: The median household income for Loop is $93,254 dollars, which means most people hold a lot of objects of value that thieves can get a hold of. As well, the area has a relatively small population, which exaggerates the theft rate. 










